---
title: "Testing"
author: "Nitay Alon"
date: "May 5, 2020"
output: html_document
---

```{r load environemt}
library(beepr)
library(dplyr)
library(tidyverse)
library(Rcpp)
```

```{r determine starting day}
covid_data <- 
full_data_for_export %>% 
  select(-var) %>% 
  as.matrix()
starting_day <- seq(40:75)
starting_day_params <- lapply(starting_day, function(i){RunFullCycle(covid_data, i)})
```

```{r devide the day into binns}
partition_parameter = 100
del = 1/partition_parameter # dt
# Set the first 100 obs to same value - can be 52,53
asympt = c()
LOGL = VAR1 = OBJ = OBJB = c()
T_final = T1 = matrix(0, ncol = 2, nrow = n)
calibration_loops = 15
hhh_upper_limit = 400
gamma_hat = beta_hat = matrix(nrow = calibration_loops, ncol = hhh_upper_limit)
NNN = floor(max(X) / 100) #What is the grid of K
alpha = 0.5
```

1. Keep $\alpha = 1$
2. K = pop
3. Best random time

A. Repeat but with hhh finds the MLE
B. Set hhh by min-log-like
C. Find the optimal value of K
D. $alpha = 0.5$

Repeat for US as well
```{r grid search for K}
K_grid_results <- lapply(1:hhh_upper_limit, function(i){
  GridSearchForK(i,X,VW,Y,alpha,Y_middle,X_middle, NNN)
})
OBJ <- sapply(K_grid_results, function(x){x$OBJ})
OBJB <- sapply(K_grid_results, function(x){x$OBJB})
```

```{r llk minimization}
plot(1:hhh_upper_limit, OBJ, type = 'l', col = 'blue', ylim = c(200,800))
lines(1:hhh_upper_limit, OBJB, type = 'l' , col = 'red')
```


Next we minimize OBJ and OBJB to find the optimal value of hhh
When we compute the CI we should consider the dimension of the data (Chi2)
```{r computing CI for K}
optimal_ks <- computeOptimalK(OBJB, OBJ, NNN, X)
```


```{r for each value of K compute the other sizes}
one_dim_params <- lapply(optimal_ks$single_dim_CI, function(k){
  InnerCalibrationLoop(k,X,VW,Y,alpha,Y_middle,X_middle)
})

one_dim_params[[2]]$beta
one_dim_params[[2]]$gamma * 100
```
beta should be $1.227*10^-4$  and gamma should be $2.58187*10^-2$

```{r validation plots}
plot(X/ 1e5, col = 'blue' , type = 'b', pch=7, ylim = c(0,2.5), panel.first = grid())
lines(VW / 1e5,col = 'blue', type = 'b', pch=13)
lines(Y / 1e5, col = 'blue', type = 'b', pch=19)

lines(x_mle[(1:length(x_mle)) %% 100 == 0] / 1e5, col = 'red', type = 'b', pch=7)
lines(vw_mle[(1:length(vw_mle)) %% 100 == 0]/ 1e5, col = 'red', type = 'b', pch=13)
lines(y_mle[(1:length(y_mle)) %% 100 == 0] / 1e5, col = 'red', type = 'b', pch=19)
```

```{r model plots}
plot(log(x[(1:length(x)) %% 100 == 0]), type = 'l', col = 'blue', ylab = 'Log cases', xlab = 'Time')
lines(log(y[(1:length(x)) %% 100 == 0]), col = 'red')
lines(log(vw[(1:length(x)) %% 100 == 0]), col = 'green')
```

```{r validation plots}
plot(x[(1:length(x)) %% 100 == 0]/ 1e5, type = 'l', col = 'blue')
lines(X/ 1e5, col = 'red')

plot(vw[(1:length(vw)) %% 100 == 0]/ 1e5, col = 'blue', type = 'l')
lines(VW / 1e5,col = 'red')

plot(y[(1:length(y)) %% 100 == 0] / 1e5, col = 'blue', type= 'l')
lines(Y / 1e5, col = 'red', type='l')
```

 Take the argmin and search for the first value where the llk increase by 10 (two-sided)

The entire code is wrapped in an external loop to test the starting day - decided by var1
Search for min(var1) * 1.5 and find, before the argmin 
Run only on days starting with x > 100 

Report $\gamma$, $\beta$, $K$, for each day and see that the starting day make no difference
$\beta ^ \frac{2}{3}$
